### 2014-11-10 Arvind.tex
stuff

Popular instruction set architectures, such as Intel x86, ARM, and PowerPC, do not guarantee sequential consistency.  That is, their behavior makes it unsound to imagine that
the different processors of a shared-memory system are running
according to simple interleaving of instructions, where each instruction executes atomically.  However, we want to emphasize that, in every weak-memory system we are aware of, \emph{the main memory still exposes atomic loads and stores}!  The weaker semantics arises only because of (1)
reordering of memory instructions by the core $P_i$ and/or (2) the
properties of the local buffers $B_i$ connected to each processor
$P_i$.  For instance, if $B_i$ were a store buffer, into which the
committed stores from an in-order processor $P_i$ are enqueued, and if
any ``subsequent'' load from the same processor $P_i$ is allowed to read and commit the latest value from the store buffer, then such a multiprocessor system is said to implement the TSO~\cite{x86tsocacm10} memory
model.

>>>>

Popular instruction set architectures, such as Intel x86, ARM, and PowerPC, do not guarantee sequential consistency.  That is, their behavior makes it unsound to imagine that
the different processors of a shared-memory system are running
according to simple interleaving of instructions, where each instruction executes atomically.  However, we want to emphasize that, in every weak-memory system we are aware of, \emph{the main memory still exposes atomic loads and stores}!  The weaker semantics arises only because of (1)
reordering of memory instructions by the core $P_i$ and/or (2) the
properties of the local buffers $B_i$ connected to each processor
$P_i$.  For instance, if $B_i$ were a store buffer, into which the
committed stores from an in-order processor $P_i$ are enqueued, and if
any ``subsequent'' load from the same processor $P_i$ is allowed to read and commit the latest value from the store buffer, then such a multiprocessor system is said to implement the TSO memory model \cite{x86tsocacm10}.

### 2014-11-10 Introduction.tex
stuff

In this paper, we introduce \emph{a new approach to modular, deductive
verification of hardware designs} at the highest level of rigor, backed by
proofs using the Coq proof assistant.  As a challenging case study, we focus on
\textbf{verifying that a particular infinite family of multicore, shared-memory
systems implements sequential consistency}.  The proof is parametrized over an
unknown number of processors connected to an arbitrary memory hierarchy with an
unknown number of caches in an unknown number of layers (e.g., L1, L2).  In our
design, processors and memory systems both independently employ intricate
optimizations that exploit opportunities for parallelism.  We are able to prove
that each of those two main components still provides strong enough guarantees
to support SC, and then we compose those theorems into a result for the full
system.  Either component may be optimized further without requiring any
changes to the implementation, specification, or proof of the other.

@@

LTSes as hardware descriptions are an established
idea \cite{HoeArvind:TRSSynthesis1, Hoe:TCAD}, and there are compilers that
convert LTSes into efficient hardware.  Our work is based on the Bluespec
language \cite{BSV:LangRef, Bluespec:TFRG}, which served as the model for the
formalism of this paper.  Bluespec models hardware components as atomic rules
of a transition system over state elements, and there is a commercial compiler
to synthesize such codes into register-transfer-level circuits (i.e., Verilog code) with
competitive performance.  Our cache-coherent memory system is directly
transliterated from a Bluespec
implementation \cite{DNA:CoherenceImplementation} used to implement a
cycle-accurate simulator for a multiprocessor PowerPC system \cite{Khan:PowerPc}. The hardware synthesized
from that implementation is rather efficient: an 8-core system with a 2-level
cache hierarchy can run 55 million instructions per second on the BEE FPGA
board \cite{bee2}. Figure \ref{both} gives an example of an LTS inference rule
transliterated directly from an actual Bluespec rule used in the PowerPC
system.  Its meaning will become clear in Section \ref{sec:cc};
for now we just note that, while Bluespec rules employ Boolean \emph{guard}
conditions paired with sequences of imperative \emph{actions} to update state,
our (equivalent) formalization follows a natural-deduction inference-rule style,
where we explain a state transition as a relation between two tuples of state
elements, using rule premises where Bluespec uses guards.  The high-level point
is that the natural-deduction style is a semantically trivial rephrasing of the
Bluespec notation, which has been demonstrated as practical for synthesizing
realistic, high-performance hardware.  Inference rules like the one in
Figure \ref{both} fit naturally with the inductive-definition mechanisms of
proof systems like Coq, providing an effective starting point for formal
verification of hardware designs.

@@

\textbf{endrule}
\end{boxedminipage}
\begin{boxedminipage}{\columnwidth}
\small
\inference[]
{\parent(c,p) & \s(c,a) < x & \wt(c,a) = \epsilon}
{\ioxe{$M_c$}
{\dt, \ch, \s,\\ \dst, \wt, \dwt,\\ \inp, \outp}
{\dt, \ch[ (c, p, \req) \coloneqq(a, \s(c, a),x)\\
\;\;\uplus \ch(c,p,\req)], \s, \dst,\\
\wt[(c,a)\coloneqq x], \dwt,\inp, \outp}{}}
\end{boxedminipage}
\caption{Bluespec rule and its corresponding LTS rule}
\label{both}
\end{figure}

>>>>

In this paper, we introduce \emph{a new approach to modular, deductive
verification of hardware designs} at the highest level of rigor, backed by
proofs using the Coq proof assistant.  As a challenging case study, we focus on
\emph{verifying that a particular infinite family of multicore, shared-memory
systems implements sequential consistency}.  The proof is parametrized over an
unknown number of processors connected to an arbitrary memory hierarchy with an
unknown number of caches in an unknown number of layers (e.g., L1, L2).  In our
design, processors and memory systems both independently employ intricate
optimizations that exploit opportunities for parallelism.  We are able to prove
that each of those two main components still provides strong enough guarantees
to support SC, and then we compose those theorems into a result for the full
system.  Either component may be optimized further without requiring any
changes to the implementation, specification, or proof of the other.

@@

LTSes as hardware descriptions are an established
idea \cite{HoeArvind:TRSSynthesis1, Hoe:TCAD}, and there are compilers that
convert LTSes into efficient hardware.  Our work is based on the Bluespec
language \cite{BSV:LangRef, Bluespec:TFRG}, which served as the model for the
formalism of this paper.  Bluespec models hardware components as atomic rules
of a transition system over state elements, and there is a commercial compiler
to synthesize such codes into register-transfer-level circuits (i.e., Verilog code) with
competitive performance.  Our cache-coherent memory system is directly
transliterated from a Bluespec
implementation \cite{DNA:CoherenceImplementation} used to implement a
cycle-accurate simulator for a cache-coherent multiprocessor PowerPC system \cite{Khan:PowerPc}. The hardware synthesized
from that implementation is rather efficient: an 8-core system with a 2-level
cache hierarchy can run 55 million instructions per second on the BEE FPGA
board \cite{bee2}.

Figure \ref{both} gives an example of an LTS inference rule
transliterated directly from an actual Bluespec rule used in the PowerPC
system.  Its meaning will become clear in Section \ref{sec:cc};
for now we just note that, while Bluespec rules employ Boolean \emph{guard}
conditions paired with sequences of imperative \emph{actions} to update state,
our (equivalent) formalization follows a natural-deduction inference-rule style,
where we explain a state transition as a relation between two tuples of state
elements, using rule premises where Bluespec uses guards.  The high-level point
is that the natural-deduction style is a semantically trivial rephrasing of the
Bluespec notation, which has been demonstrated as practical for synthesizing
realistic, high-performance hardware.  Inference rules like the one in
Figure \ref{both} fit naturally with the inductive-definition mechanisms of
proof systems like Coq, providing an effective starting point for formal
verification of hardware designs.

@@

\textbf{endrule}
\end{boxedminipage}
\begin{boxedminipage}{\columnwidth}
\small
\inference[]
{\parent(c,p) & \s(c,a) < x & \wt(c,a) = \epsilon}
{\hspace{-.4cm}\ioxe{$M_c$}
{\hspace{-.2cm}\dt, \ch, \s,\hspace{-.2cm}\\
\hspace{-.2cm}\dst, \wt, \dwt,\hspace{-.2cm}\\
\hspace{-.2cm}\inp, \outp\hspace{-.2cm}}
{\hspace{-.2cm}\dt, \ch[ (c, p, \req) \coloneqq(a, \s(c, a),x)\hspace{-.2cm}\\
\;\;\;\;\;\;\uplus \ch(c,p,\req)], \s, \dst,\hspace{-.2cm}\\
\hspace{-.2cm}\wt[(c,a)\coloneqq x], \dwt,\inp, \outp\hspace{-.2cm}}{}\hspace{-.4cm}}
\end{boxedminipage}
\caption{Bluespec rule and its corresponding LTS rule}
\label{both}
\end{figure}
